Hydras \& Co.

### 2023-12-08 doc/chapter-primrec.tex
Version for Coq <= 8.17.1  (heavily uses Coq.Sets and Coq.Vector)

We may want to prove properties shared with all the set of
primitive function, for instance that any primitive recursive function is representable by a predicate in Peano Arithmetics (see Chap~›\vref{chap:PRRep}).
In this chapter, we will prove by induction that every primitive recursive function has a given property which is not satisfied by  Ackermann function, and deduce that this function is not PR.
We show how to adapt the classic proof (see for instance~\cite{planetmath}) to the constraints of \gallina. We hope this formal proof 
 is a nice opportunity to explore
the treatment of primitive recursive functions by R. O'Connor,
and to play with dependent types.

First, let us define Ackermann function in \coq, then look at a few properties of this function.

@@

We preferred to define a variant which uses explicitly
 the functional \texttt{iterate},
where (\texttt{iterate\,$f$\,$n$})
is the $n$-th iteration of $f$\,\footnote{Please not confuse with \texttt{primRec.iterate}, which is monomorphic and does not share the same order of arguments.}. It makes it possible to apply a few lemmas proved in 
\href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}, for instance about the monotony of the $n$-th iterate of a given function.

>>>>

Let us consider the following theorem (see for instance ~\cite{planetmath}).

\begin{quote}
   There exists at least a total arithmetic function, \emph{e.g.} the \emph{Ackermann function},  which is not primitive recursive.
\end{quote}

We can prove this theorem in three successive steps:

\begin{itemize}
\item Define Ackermann function in \gallina.
\item Define and prove a property shared by any primitive recursive functions.
\item Prove that  Ackermann function does not satisfy this property.
\end{itemize}

We show how to adapt the classic proof (see for instance~\cite{planetmath}) to the constraints of \gallina. We hope this formal proof 
 is a nice opportunity to explore
the treatment of primitive recursive functions by R. O'Connor,
and to play with dependent types.

@@

We preferred to define a variant which uses explicitly
 the functional \texttt{iterate},
where (\texttt{iterate\,$f$\,$n$})
is the $n$-th iteration of $f$\,\footnote{Please do not mistake
  \texttt{iterate} (\emph{i.e.} \texttt{Prelude.iterate}) for
  the monomorphic \texttt{primRec.iterate}, which does not share the same order of arguments.}. It makes it possible to apply a few lemmas proved in 
\href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}, for instance about the monotony of the $n$-th iterate of a given function.
